---
title: "Course_Project"
author: "suveyda"
date: "January 31, 2017"
output: html_document
---

# Course Project

## Getting and Cleaning Data by Johns Hopkins University

### Introduction

One of the most exciting areas in all of data science right now is wearable computing. Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The purpose of this project is to demonstrate our ability to collect, work with, and clean a data set collected from the accelerometers from the Samsung Galaxy S smartphone. The goal is to prepare tidy data that can be used for later analysis.

### Data

Human Activity Recognition database built from the recordings of 30 subjects performing activities of daily living (ADL) while carrying a waist-mounted smartphone with embedded inertial sensors. Data is available at the UCI Machine Learning Repository as [Human Activity Recognition Using Smartphones Data Set](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones).

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

#### Attribute Information:

For each record in the dataset it is provided:

* Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration. 
* Triaxial Angular velocity from the gyroscope. 
* A 561-feature vector with time and frequency domain variables. 
* Its activity label. 
* An identifier of the subject who carried out the experiment.

Let's first read the training and test set along with their labels:

```{r}
X_train <- read.table("./UCI HAR Dataset/train/X_train.txt", header = FALSE)
X_test <- read.table("./UCI HAR Dataset/test/X_test.txt", header = FALSE)
y_train <- read.table("./UCI HAR Dataset/train/y_train.txt", header = FALSE)
y_test <- read.table("./UCI HAR Dataset/test/y_test.txt", header = FALSE)
```

Now we read the subject id for both training and test set observations:

```{r}
subject_train <- read.table("./UCI HAR Dataset/train/subject_train.txt", header = FALSE)
subject_test <- read.table("./UCI HAR Dataset/test/subject_test.txt", header = FALSE)
```

We also read the list of all features to be used as column names:

```{r}
features <- read.table("./UCI HAR Dataset/features.txt", header = FALSE, stringsAsFactors=FALSE)
```

We merge the training and the test sets to create one data set, and appropriately label each the data set with descriptive variable names:

```{r}
X <- rbind(X_train, X_test)
names(X) <- features[,2]

y <- rbind(y_train, y_test)
names(y) <- "activity"

subject <- rbind(subject_train, subject_test)
names(subject) <- "subject"
```

Now we extract only the measurements on the mean and standard deviation for each measurement:

```{r}
sub_X <- subset(X, select = grep("mean\\(\\)|std\\(\\)", names(X)))
```

We also replace the activity labels with descriptive activity names:

```{r}
y <- lapply(y, function(x) gsub("1", "WALKING", x))
y <- lapply(y, function(x) gsub("2", "WALKING_UPSTAIRS", x))
y <- lapply(y, function(x) gsub("3", "WALKING_DOWNSTAIRS", x))
y <- lapply(y, function(x) gsub("4", "SITTING", x))
y <- lapply(y, function(x) gsub("5", "STANDING", x))
y <- lapply(y, function(x) gsub("6", "LAYING", x))
```

Now we combine subject information, activity labels, and measurements data in one big data frame:

```{r}
data <- cbind(subject, y, sub_X)
```

Finally, we create a second, independent tidy data set with the average of each variable for each activity and each subject. At the end, we write this tidy data set into a file.

```{r}
data.mean <- aggregate(data[,-c(1,2)], list(subject=data$subject, activity=data$activity), mean)
write.table(data.mean, "tidy.txt", row.names = FALSE, quote = FALSE)
```

Here you can see the tidy data set:

```{r}
str(data.mean)
```

Now, this data is ready to be used in further analysis.
